\ZpL: a p-adic precision package

### 2018-01-30 ZpL/ZpL.tex

\newcommand{\sage}{\textsc{SageMath}\xspace}
\newcommand{\ZpCR}{\text{\rm \tt ZpCR}\xspace}
\newcommand{\ZpFP}{\text{\rm \tt ZpFP}\xspace}
\newcommand{\ZpL}{\text{\rm \tt ZpL}\xspace}
\newcommand{\ZpLC}{\text{\rm \tt ZpLC}\xspace}
\newcommand{\ZpLF}{\text{\rm \tt ZpLF}\xspace}

\newcommand{\card}{\text{Card}\:}
\newcommand{\ind}{\text{ind}}
\newcommand{\coind}{\text{coind}}

@@

\noindent \textbf{Creating a variable.}

\ZpLC: $O\bigg(\sum_{i=1}^n \ind_t(\ttv_i)\bigg) \subset 
O\big(n \cdot \card \calV_t\big)$

\ZpLF:

@@

As we can see, some loss in precision occurs in the
Buchberger algorithm and is avoided thanks to \ZpL.
Had we used \cZpFP , many padded zeroes or one would have appeared.

>>>>

\newcommand{\sage}{\textsc{SageMath}\xspace}
\newcommand{\ZpCR}{\text{\rm \tt ZpCR}\xspace}
\newcommand{\ZpFP}{\text{\rm \tt ZpFP}\xspace}
\newcommand{\ZpL}{\text{\rm \tt ZpL}\xspace}
\newcommand{\ZpLC}{\text{\rm \tt ZpLC}\xspace}
\newcommand{\ZpLF}{\text{\rm \tt ZpLF}\xspace}
\newcommand{\ZpFP}{\text{\rm \tt ZpFP}\xspace}

\newcommand{\card}{\text{Card}\:}
\newcommand{\ind}{\text{ind}}
\newcommand{\coind}{\text{coind}}
\newcommand{\hgt}{\text{hgt}}
\newcommand{\cohgt}{\text{cohgt}}
\renewcommand{\in}{\text{in}}
\newcommand{\out}{\text{out}}

@@

\noindent \textbf{Creating a variable.}
%
With the notations of \S \ref{ssec:trackprec}, it is obvious that 
creating a new variable \ttw requires:
$$O\bigg(\sum_{i=1}^n \hgt_i(\ttv_i)\bigg) \subset O(n \: c_t)$$ 
operations in $\Qp$. Here, we recall that $n$ is the arity of the 
operation defining \ttw. In most cases it is $2$; thus the above
complexity reduces to $O(c_t)$.

In the \ZpLF context, $c_t$ counts the number of user-defined variables. 
It is then expected to be constant (roughly equal to the size of the 
input) while running a given algorithm.

On the contrary, in the \ZpLC context, $c_t$ counts the number of
variables which are alive at time $t$. It is no longer expected to
be constant but evolves continuously when the algorithm runs.
%
\begin{figure}
% Characteristic polynomial
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent\hfill%
\begin{tabular}{|l|r|r|r|r|r|}
\hline
\textbf{Dimension} &
 2 &   5 &   10 &    20 &      50 \\
\hline
\textbf{Total} &
35 & 424 & 5\:539 & 83\:369 & 3\:170\:657 \\
\hline
\textbf{Simult.} &
17 &  65 &  225 &   845 &    5\:101 \\
\hline
\end{tabular}%
\hfill\null

\vspace{1mm}

\noindent\hfill%
{\small Computation of characteristic polynomial}%
\hfill\null

\bigskip

% Euclidean algorithm

\noindent\hfill%
\begin{tabular}{|l|r|r|r|r|r|r|}
\hline
\textbf{Degree} &
 2 &   5 &   10 &   20 &   50 &   100 \\
\hline
\textbf{Total} &
54 & 130 &  332 & 1\:036 & 4\:110 & 10\:578 \\ 
\hline
\textbf{Simult.} &
18 &  31 &   56 &  106 &  256 &   507 \\
\hline
\end{tabular}%
\hfill\null

\vspace{1mm}

\noindent\hfill%
{\small Naive Euclidean algorithm}%
\hfill\null

\caption{Numbers of involved variables}
\label{fig:variables}
\end{figure}
%
The tables of Figure~\ref{fig:variables} show the total number of 
created variables (which reflects the complexity) together with the 
maximum number of variables involved at the same time (which reflects 
the memory occupation) while executing two basic computations.
The first one is the computation of the characteristic polynomial of a 
square matrix by the default algorithm used by \sage for $p$-adic fields 
(which is a division-free algorithm of quartic complexity) while the
second one is the computation of the gdc of two polynomials using a
naive Euclidean algorithm (of quadratic complexity). We can observe 
that, for both of them, the memory occupation is roughly equal to the 
square root of the complexity.

@@

As we can see, some loss in precision occurs in the
Buchberger algorithm and is avoided thanks to \ZpL.
Had we used \ZpFP, many padded zeroes or one would have appeared.
